Supporting Information for "Geometric models overestimate lake depth due to imperfect slope measurement"

### 2020-11-10 manuscript/si_template_2019.tex
relabel appendix figs

\begin{figure}
      \begin{center}
        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
      \end{center}
      \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right). Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
    \end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figs1}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figs4}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope. A dashed 1:1 line is shown for comparison. Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figs5}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
      \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figs6}
      \end{figure}

\begin{figure}
      \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
            \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product. The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figs7}
      \end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figs3}
      \end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product. For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figs2}
      \end{figure}

@@

\begin{table}
\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
\centering
\setlength\tabcolsep{1.5pt} % default value: 6pt
\begin{tabular}{lllll}
  \hline
  Variable & Median & Q25 & Q75 & n\\
  \hline
  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\
  \hline
  \end{tabular}
\end{table}

@@

More Information and Advice:

@@

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

@@

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

>>>>

\begin{figure}
      \begin{center}
        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
      \end{center}
      \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right). Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{figS1}
    \end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figS2}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS3}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope. A dashed 1:1 line is shown for comparison. Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS4}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
      \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS5}
      \end{figure}

\begin{figure}
      \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
            \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product. The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS6}
      \end{figure}

\clearpage
\begin{figure}
      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS7}
      \end{figure}

@@

\begin{figure}
      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product. For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS8}
      \end{figure}

@@

\begin{table}
\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{tableS1}
\centering
\setlength\tabcolsep{1.5pt} % default value: 6pt
\begin{tabular}{lllll}
  \hline
  Variable & Median & Q25 & Q75 & n\\
  \hline
  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\
  \hline
  \end{tabular}
\end{table}

@@

% More Information and Advice:

@@

% Math coded inside display math mode \[ ...\]
%  will not be numbered, e.g.,:
%  \[ x^2=y^2 + z^2\]
%
%  Math coded inside \begin{equation} and \end{equation} will
%  be automatically numbered, e.g.,:
%  \begin{equation}
%  x^2=y^2 + z^2
%  \end{equation}

@@

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
% \begin{eqnarray}
%   x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
%         && + (y - y_{0}) \sin \Theta  \nonumber \\
%   y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
%         && + (y - y_{0}) \cos \Theta.
% \end{eqnarray}
